---
title: "First analysis reading gains"
author: "Samuel Merk, Sarah Bez"
format: 
  html:
    toc: true
    theme: 
      light: flatly
      dark: darkly
    lightbox: true
editor_options: 
  chunk_output_type: console
---


## Variables

* pid = pseudonym kind
* elfe_pre = Elfe score prä 
* elfe_post = elfe score post
* treat = Variable für treatment, 0 = kein t, 1 = treatment 
* treat_drop = Variable für dropout oder späteres Hinzukommen beim treatment, 0 = nein, 1 = ja 

# Overview

## Kohorte 1 MA und KA

### Descriptive Overview

```{r}
#| message: false

library(skimr)
library(ggforce)
library(hrbrthemes)
library(MatchThem)
library(mice)
library(cobalt)
library(naniar)
library(patchwork)
library(brms)
library(bayesplot)
library(bayestestR)
library(sjPlot)
library(effectsize)
library(tidyverse)

data_k1 <- read_csv("data/data_ELFE_KA_k1.csv") |> 
  column_to_rownames("...1") |> 
  mutate(dummy_foerderung = as.factor(treat),
         school = as.factor(substr(pid, 1, 1)),
         class = as.factor(substr(pid, 1, 2)),
         gender_pre = as.factor(gender_pre),
         migration_pre = as.factor(migration_pre),
         language_pre = as.factor(language_pre),
         uni = "KA") |> 
  rename(elfe_t_pre = elfe_pre,
         elfe_t_post = elfe_post) |> 
  as_tibble() |> 
  full_join(
    read_csv("data/data_ELFE_MA_k1.csv") |> 
  column_to_rownames("...1") |> 
  mutate(dummy_foerderung = as.factor(dummy_foerderung),
         school = as.factor(substr(pid, 1, 3)),
         class = as.factor(substr(pid, 1, 5)),
         gender_pre = as.factor(gender_pre),
         migration_pre = as.factor(migration_pre),
         language_pre = as.factor(language_pre),
        #elfe_t_pre = elfe_pre,
         #elfe_t_post = elfe_post,
         pid = as.character(pid),
         uni = "MA",
         klassenstufe = case_when(class == 60011 ~ 3, # to avoid missings in _pre und _post
                                       class == 60012 ~ 2, 
                                       class == 60111 ~ 4, 
                                       class == 60112 ~ 4, 
                                       class == 60211 ~ 3, 
                                       class == 60212 ~ 3))%>%
      select(-klassenstufe_post)%>%
      select(-klassenstufe.y)%>%
    select(-klassenstufe.x)%>%
    as_tibble()
  )
    

skim(data_k1)


```

### Demographics

```{r}
#classes
data_k1 |> 
  filter(uni == "KA") |> 
  pull(pid) |> 
  substr(1,2) |> 
  table()

data_k1 |> 
  filter(uni == "MA") |> 
  pull(pid) |> 
  substr(1,5) |> 
  table()

unique(data_k1$pid)

sum(as.numeric(data_k1$dummy_foerderung) -1)
```


## Kohorte 2 KA 

### Descriptive overview
```{r}
data_k2_ka <- read_csv("data/data_ELFE_KA_k2.csv") |> 
  column_to_rownames("...1") |> 
  mutate(dummy_foerderung = as.factor(treat),
         school = as.factor(substr(pid, 1, 1)),
         class = as.factor(substr(pid, 1, 2)),
         gender_pre = as.factor(gender_pre),
         migration_pre = as.factor(migration_pre),
         language_pre = as.factor(language_pre),
         uni = "KA")%>%
  rename(elfe_t_pre = elfe_pre,
         elfe_t_post = elfe_post)%>%
  select(-Uni)

skim(data_k2_ka)

```


### Demographics

```{r}
#classes
data_k2_ka |> 
  filter(uni == "KA") |> 
  pull(pid) |> 
  substr(1,2) |> 
  table()

unique(data_k2_ka$pid)

sum(as.numeric(data_k2_ka$dummy_foerderung) -1)
```

# Retestcorrelations

## Retestcorrelations k1

```{r}
library(corrgram)
corrgram(data_k1 |> 
           select(elfe_t_pre, elfe_t_post,
                  tv_sum_score_pre, tv_sum_score_post,
                  sv_sum_score_pre, sv_sum_score_post, 
                  wv_sum_score_pre, wv_sum_score_post,
                  tv_speed_t_pre, tv_speed_t_post,
                  sv_speed_t_pre, sv_speed_t_post,
                  wv_speed_t_pre, wv_speed_t_post
                  ),
         upper.panel = panel.pie,
         lower.panel = panel.cor
)

corrgram(data_k1 |> 
           filter(uni == "KA") |> 
           select(elfe_t_pre, elfe_t_post,
                  tv_sum_score_pre, tv_sum_score_post,
                  sv_sum_score_pre, sv_sum_score_post, 
                  wv_sum_score_pre, wv_sum_score_post,
                  tv_speed_t_pre, tv_speed_t_post,
                  sv_speed_t_pre, sv_speed_t_post,
                  wv_speed_t_pre, wv_speed_t_post
                  ),
         upper.panel = panel.pie,
         lower.panel = panel.cor
)

corrgram(data_k1 |> 
           filter(uni == "MA") |> 
           select(elfe_t_pre, elfe_t_post,
                  tv_sum_score_pre, tv_sum_score_post,
                  sv_sum_score_pre, sv_sum_score_post, 
                  wv_sum_score_pre, wv_sum_score_post,
                  tv_speed_t_pre, tv_speed_t_post,
                  sv_speed_t_pre, sv_speed_t_post,
                  wv_speed_t_pre, wv_speed_t_post
                  ),
         upper.panel = panel.pie,
        lower.panel = panel.cor
)
```

## Retestcorrelations k2 KA

```{r}
library(corrgram)
corrgram(data_k2_ka |> 
           select(elfe_t_pre, elfe_t_post,
                  tv_sum_score_pre, tv_sum_score_post,
                  sv_sum_score_pre, sv_sum_score_post, 
                  wv_sum_score_pre, wv_sum_score_post,
                  tv_speed_t_pre, tv_speed_t_post,
                  sv_speed_t_pre, sv_speed_t_post,
                  wv_speed_t_pre, wv_speed_t_post
                  ),
         upper.panel = panel.pie,
         lower.panel = panel.cor
)

corrgram(data_k2_ka |> 
           select(elfe_t_post,
                   tv_sum_score_post,
                   sv_sum_score_post, 
                   wv_sum_score_post,
                   tv_speed_t_post,
                   sv_speed_t_post,
                   wv_speed_t_post
                  ),
         upper.panel = panel.pie,
         lower.panel = panel.cor
)

corrgram(data_k2_ka |> 
           select(elfe_t_pre, 
                  tv_sum_score_pre, 
                  sv_sum_score_pre, 
                  wv_sum_score_pre, 
                  tv_speed_t_pre,
                  sv_speed_t_pre,
                  wv_speed_t_pre
                  ),
         upper.panel = panel.pie,
         lower.panel = panel.cor
)
```

# All data (except k2 MA)

```{r}
data <- full_join(data_k1, data_k2_ka) %>% 
  mutate(gain_t = elfe_t_post - elfe_t_pre,
         gain_sum_wv = wv_sum_score_post - wv_sum_score_pre,
         gain_sum_sv = sv_sum_score_post - sv_sum_score_pre,
         gain_sum_tv = tv_sum_score_post - tv_sum_score_pre)

skim(data)
```

## Demographics

```{r}
#classes
unique(data$class)

#students
unique(data$pid)

#Förderkinder
sum(as.numeric(data$dummy_foerderung) -1)
```

## Missingness

### Plotting Missingness

```{r}
data_to_imp <- 
  data |>
  select(elfe_t_pre,
         gender_pre,
         migration_pre,
         language_pre,
         elfe_t_post,
         wv_sum_score_pre,
         sv_sum_score_pre,
         tv_sum_score_pre,
         wv_sum_score_post,
         sv_sum_score_post,
         tv_sum_score_post,
         class) |>
  tibble() 

vis_miss(data_to_imp)

gg_miss_upset(data_to_imp,
              nsets = 10)

ggplot(data_to_imp,
       aes(x = elfe_t_pre,
           y = elfe_t_post)) +
 geom_miss_point() +
 theme_modern_rc()
   
ggplot(data_to_imp,
       aes(x = wv_sum_score_pre,
           y = wv_sum_score_post)) +
 geom_miss_point() +
 theme_modern_rc()

ggplot(data_to_imp,
       aes(x = wv_sum_score_pre,
           y = wv_sum_score_post)) +
 geom_miss_point() +
 theme_modern_rc()

ggplot(data_to_imp,
       aes(x = tv_sum_score_pre,
           y = tv_sum_score_post)) +
 geom_miss_point(alpha = .3) +
 theme_modern_rc()
```



## Differences in gain
### Data Wrangling

```{r}
data_long <- 
  data |> 
  pivot_longer(names_to = "variable",
               values_to = "value",
               cols = c(contains("pre"), contains("post")),
               values_transform = list(value = as.character)) |> 
  mutate(time = ifelse(str_detect(variable, "pre"), "pre", "post"),
         variable2 = ifelse(time == "pre",
                            str_sub(variable,-100,  -5),
                            str_sub(variable,-100,  -6))) %>% 
  select(-variable) %>% 
  pivot_wider(names_from = variable2,
              values_from = value) %>% 
  mutate(time = factor(time, levels = c("pre", "post"))) %>% 
  mutate(across(contains("sum_score"), as.numeric),
         across(contains("speed"), as.numeric),
         across(contains("_t"), as.numeric))
```





## Gain-t linear Regression
### Homoskedastic Linear Regression 
```{r}
ggplot(data,
       aes(elfe_t_pre, elfe_t_post, 
           color = dummy_foerderung)) + 
  geom_jitter() +
  stat_smooth()

ggplot(data,
       aes(dummy_foerderung, gain_t)) + 
  geom_sina(aes(color = elfe_t_pre)) + 
  stat_summary(fun.data = mean_sdl, 
               geom = "errorbar",
               color = "lightgrey",
               fun.args = list(mult = 1),
               width = 0) +
  stat_summary(fun.data = mean_sdl, 
               geom = "point",
               color = "lightgrey") +
  theme_modern_rc() +
  scale_color_continuous(type = "viridis")

# Single Level Model
mod_gain_t_homo <- 
  brm(bf(gain_t ~ dummy_foerderung),
      family = student(),
      data = data)

pp_check(mod_gain_t_homo)
plot(mod_gain_t_homo)
summary(mod_gain_t_homo)
hypothesis(mod_gain_t_homo, "dummy_foerderung1 > 0" )

# Fixed Effects Model
mod_gain_t_homo_fe <- 
  brm(bf(gain_t ~ dummy_foerderung + class),
      family = student(),
      iter = 40000,
      cores = 4,
      data = data)

pp_check(mod_gain_t_homo_fe)
plot(mod_gain_t_homo_fe)
summary(mod_gain_t_homo_fe)
hypothesis(mod_gain_t_homo_fe, "dummy_foerderung1 > 0" )

# Random Effects Model
mod_gain_t_homo_re <- 
  brm(bf(gain_t ~ dummy_foerderung + (1|class)),
      family = student(),
      iter = 40000,
      cores = 4,
      data = data)

pp_check(mod_gain_t_homo_re)
plot(mod_gain_t_homo_re)
summary(mod_gain_t_homo_re)
hypothesis(mod_gain_t_homo_re, "dummy_foerderung1 > 0" )

tab_model(mod_gain_t_homo, mod_gain_t_homo_fe, mod_gain_t_homo_re)


# Effect Size
cliffs_delta(gain_t ~ dummy_foerderung, data = data)
```

### Heteroskedastic Linear Regression (single level)
```{r}
mod_gain_t_hetero <- 
  brm(bf(gain_t ~ dummy_foerderung,
         sigma ~ dummy_foerderung),
      family = student(),
      data = data |> 
         mutate(gain_t = elfe_t_post - elfe_t_pre))

pp_check(mod_gain_t_hetero)
plot(mod_gain_t_hetero)

summary(mod_gain_t_hetero)

pd(mod_gain_t_hetero)
plot(pd(mod_gain_t_hetero))
rope(mod_gain_t_hetero, range = c(-.1, .1))
```

### Comparison to Homoscedastic Model
```{r}
loo_compare(add_criterion(mod_gain_t_homo, "loo"), 
            add_criterion(mod_gain_t_hetero, "loo"))
```

Aus Parsimonitätsgründen Homoskedastizität annehmen!



## Gain sumscore_wv
```{r}
ggplot(data,
       aes(wv_sum_score_pre, wv_sum_score_post, 
           color = dummy_foerderung)) + 
  geom_jitter() +
  stat_smooth()

ggplot(data,
       aes(dummy_foerderung, gain_sum_wv)) + 
  geom_sina(aes(color = wv_sum_score_pre)) + 
  stat_summary(fun.data = mean_sdl, 
               geom = "errorbar",
               color = "lightgrey",
               fun.args = list(mult = 1),
               width = 0) +
  stat_summary(fun.data = mean_sdl, 
               geom = "point",
               color = "lightgrey") +
  theme_modern_rc() +
  scale_color_continuous(type = "viridis")

# Single Level Model
mod_gain_wv <- 
  brm(bf(gain_sum_wv ~ dummy_foerderung),
      family = student(),
      iter = 40000,
      cores = 4,
      data = data)

pp_check(mod_gain_wv)
plot(mod_gain_wv)
summary(mod_gain_wv)
hypothesis(mod_gain_wv, "dummy_foerderung1 > 0" )

# Fixed Effects Model
mod_gain_wv_homo_fe <- 
  brm(bf(gain_sum_wv ~ dummy_foerderung + class),
      family = student(),
      iter = 40000,
      cores = 4,
      data = data)

pp_check(mod_gain_wv_homo_fe)
plot(mod_gain_wv_homo_fe)
summary(mod_gain_wv_homo_fe)
hypothesis(mod_gain_wv_homo_fe, "dummy_foerderung1 > 0" )

# Random Effects Model
mod_gain_wv_homo_re <- 
  brm(bf(gain_sum_wv ~ dummy_foerderung + (1|class)),
      family = student(),
      iter = 40000,
      cores = 4,
      data = data)

pp_check(mod_gain_wv_homo_re)
plot(mod_gain_wv_homo_re)
summary(mod_gain_wv_homo_re)
hypothesis(mod_gain_wv_homo_re, "dummy_foerderung1 > 0" )

# Compare Models
tab_model(mod_gain_wv, mod_gain_wv_homo_fe, mod_gain_wv_homo_re)

# Effect Size
cliffs_delta(gain_sum_wv ~ dummy_foerderung, data = data)
```


## Gain sumscore_sv 
```{r}
ggplot(data,
       aes(sv_sum_score_pre, sv_sum_score_post, 
           color = dummy_foerderung)) + 
  geom_jitter() +
  stat_smooth()

ggplot(data,
       aes(dummy_foerderung, gain_sum_sv)) + 
  geom_sina(aes(color = sv_sum_score_pre)) + 
  stat_summary(fun.data = mean_sdl, 
               geom = "errorbar",
               color = "lightgrey",
               fun.args = list(mult = 1),
               width = 0) +
  stat_summary(fun.data = mean_sdl, 
               geom = "point",
               color = "lightgrey") +
  theme_modern_rc() +
  scale_color_continuous(type = "viridis")

# Single Level Model
mod_gain_sv <- 
  brm(bf(gain_sum_sv ~ dummy_foerderung),
      family = student(),
      iter = 40000,
      cores = 4,
      data = data)

pp_check(mod_gain_sv)
plot(mod_gain_sv)
summary(mod_gain_sv)
hypothesis(mod_gain_sv, "dummy_foerderung1 > 0" )

# Fixed Effects Model
mod_gain_sv_homo_fe <- 
  brm(bf(gain_sum_sv ~ dummy_foerderung + class),
      family = student(),
      iter = 40000,
      cores = 4,
      data = data)

pp_check(mod_gain_sv_homo_fe)
plot(mod_gain_sv_homo_fe)
summary(mod_gain_sv_homo_fe)
hypothesis(mod_gain_sv_homo_fe, "dummy_foerderung1 > 0" )

# Random Effects Model
mod_gain_sv_homo_re <- 
  brm(bf(gain_sum_sv ~ dummy_foerderung + (1|class)),
      family = student(),
      iter = 40000,
      cores = 4,
      data = data)

pp_check(mod_gain_sv_homo_re)
plot(mod_gain_sv_homo_re)
summary(mod_gain_sv_homo_re)
hypothesis(mod_gain_sv_homo_re, "dummy_foerderung1 > 0" )

# Compare Models
tab_model(mod_gain_sv, mod_gain_sv_homo_fe, mod_gain_sv_homo_re)

# Effect Size
cliffs_delta(gain_sum_sv ~ dummy_foerderung, data = data)
```


## Gain sumscore_tv
```{r}
ggplot(data,
       aes(tv_sum_score_pre, tv_sum_score_post, 
           color = dummy_foerderung)) + 
  geom_jitter() +
  stat_smooth()

ggplot(data,
       aes(dummy_foerderung, gain_sum_tv)) + 
  geom_sina(aes(color = tv_sum_score_pre)) + 
  stat_summary(fun.data = mean_sdl, 
               geom = "errorbar",
               color = "lightgrey",
               fun.args = list(mult = 1),
               width = 0) +
  stat_summary(fun.data = mean_sdl, 
               geom = "point",
               color = "lightgrey") +
  theme_modern_rc() +
  scale_color_continuous(type = "viridis")

# Single Level Model
mod_gain_tv <- 
  brm(bf(gain_sum_tv ~ dummy_foerderung),
      family = student(),
      iter = 40000,
      cores = 4,
      data = data)

pp_check(mod_gain_tv)
plot(mod_gain_tv)
summary(mod_gain_tv)
hypothesis(mod_gain_tv, "dummy_foerderung1 > 0" )

# Fixed Effects Model
mod_gain_tv_homo_fe <- 
  brm(bf(gain_sum_tv ~ dummy_foerderung + class),
      family = student(),
      iter = 40000,
      cores = 4,
      data = data)

pp_check(mod_gain_tv_homo_fe)
plot(mod_gain_tv_homo_fe)
summary(mod_gain_tv_homo_fe)
hypothesis(mod_gain_tv_homo_fe, "dummy_foerderung1 > 0" )

# Random Effects Model
mod_gain_tv_homo_re <- 
  brm(bf(gain_sum_tv ~ dummy_foerderung + (1|class)),
      family = student(),
      iter = 40000,
      cores = 4,
      data = data)

pp_check(mod_gain_tv_homo_re)
plot(mod_gain_tv_homo_re)
summary(mod_gain_tv_homo_re)
hypothesis(mod_gain_tv_homo_re, "dummy_foerderung1 > 0" )

# Compare Models
tab_model(mod_gain_tv, mod_gain_tv_homo_fe, mod_gain_tv_homo_re)

# Effect Size
cliffs_delta(gain_sum_tv ~ dummy_foerderung, data = data)
```
